---
title: "Data 607 Week 5"
author: "Maryluz Cruz"
date: "9/27/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Add the libraries required to do this assignment
For this assignment we need tidyr and dplyr

```{r}
require(tidyr)
```

```{r}
require(dplyr)
```

```{r}
require(tidyverse)
```

## (1) Create a .CSV file (or optionally, a MySQL database!) that includes all of the information above. You’re encouraged to use a “wide” structure similar to how the information appears above, so that you can practice tidying and transformations as described below.

This DataSet was created in Excel and uploaded into my GitHub


```{r}
airportdata <-   read.csv("https://raw.githubusercontent.com/Luz917/data607Week5/master/airport%20data.csv",stringsAsFactors = FALSE)
str(airportdata)
```



```{r}
summary(airportdata)
```

```{r}
airportdata
```

## Prepare the Table 

Before we begin to use tidyr and dplyr we need to do a feew thingsa to set up the table.
-Name all of the Columns 
-Fill in the empty spaces in the first column with the name of each airport
-Then remove the third row 

```{r}
colnames(airportdata)<-c("Airport", "FlightStatus","Los.Angelos","Phoenix","San.Diego","San.Francisco","Seattle")
airportdata [2,1]<-"ALASKA"
airportdata [5,1]<-"AM WEST"
airportdata = airportdata[-c(3) , ]
airportdata
                          
```


## Tidyr

(2) Read the information from your .CSV file into R, and use tidyr and dplyr as needed to tidy and transform your data.

We begin the transformation, we begin by putting states into a column by using pivot_longer. (source is on the bottom)

```{r}
airportdata <- pivot_longer(airportdata, cols = -c("Airport","FlightStatus"), names_to = "States")
airportdata
```

As you can see we end up with 20 rows


Next we use pivit_wider so we can see On Time and Delayed Status by Airport and City

```{r}
airportdata %>%
  pivot_wider(names_from = FlightStatus, values_from = value) %>%
  arrange(States)
```
Now there are 10 Rows


## Data Analysis with dplyr

(3) Perform analysis to compare the arrival delays for the two airlines.

## Sum of Flights by Airport

```{r}
airportdata %>%
  group_by(Airport)%>%
  summarise(AirportSum= sum(value))

```
AM West has more flights then Alaska


## Sum of Flights By State

```{r}
airportdata %>%
  group_by(States) %>%
  summarise(SumFlights= sum(value))
```
Phoenix has the most flights by state

## Sum of Flight Status by Airport

```{r}
airportdata %>%
  group_by(Airport,FlightStatus) %>%
  summarise(SumofFlights= sum(value))
```
AM West has the highest delay rate, but they also have a high number of On Time flights

## Average Number of Flights 

```{r}
airportdata %>%
  group_by(Airport,FlightStatus) %>%
  summarise(AvgFlights= mean(value))
```
AM West has the highest average number of Flights


## Max Number of Flights by State

```{r}
airportdata %>%
  group_by(Airport,States) %>%
  summarise(MaxFlights= max(value))

```
Phoenix has the Max Number of Flights 

## Max Number of Flights by Airport
```{r}
airportdata %>%
  group_by(Airport) %>%
  summarise(MaxFlights= max(value))

```
AM West has the Max Number of Flights 


## Percent Rank
```{r}
airportdata %>%
  mutate(Flights= percent_rank(value)) %>%
  arrange(States)
```
Seattle had the highest Percent Rank

## Conclusion
We can tell by the Analysis that AM West gets more flights in comparison to Alaska. Phoenix has the highest number of flights out of all of the States. 

## Sources
1. How to reshape a dataframe from wide to long or long to wide format using tidyr - pivot_longer() pivot_wider()
https://www.programmingwithr.com/how-to-reshape-a-dataframe-from-wide-to-long-or-long-to-wide-format/

2.Hands-on dplyr tutorial for faster data manipulation in R
https://www.dataschool.io/dplyr-tutorial-for-faster-data-manipulation-in-r/   

3.How do I delete rows in a data frame?
https://stackoverflow.com/questions/12328056/how-do-i-delete-rows-in-a-data-frame
